\LARGE \bf ***

### 2020-07-12 icra-2021.tex
finished model, but final proof read needed (still)

%% math package
\usepackage[cmex10]{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{arydshln}

@@

Let us define a generic continuously differentiable function $\varphi:\mathbb{R}^2\rightarrow\mathbb{R}$. The desired trajectory $\mathcal{P}$ can be hence defined:
\begin{equation}\label{eqq:2}
  \mathcal{P}:=\{\mathbf{p}:\|\varphi(\mathbf{p})\|<c+\varepsilon\},
\end{equation}
as the set of points that follows the trajectory $\varphi(\mathbf{p})=c$ within a given $\varepsilon\in\mathbb{R}^{\geq 0}$.

Given a set of discrete values $c_i$, the space relative to $\mathcal{O}_W$ can be covered by all the trajectories $\varphi(\mathbf{p})=c_i$ within specific boundaries $c_m\leq c_i\leq c_M$.

The concept of a trajectory is used later to design a controller that selects $c_i$ value with the highest energy value under the energy budget constraints defined in the next subsection. Here we design a guidance action $u$ that allows following such trajectory. The guidance is derived using the vector field approach presented in~\cite{de2017guidance}. The algorithm heads to $\mathcal{P}$ using a vector field based on the minimization of the norm in Equation~(\ref{eqq:2}).

Let us define $\varPhi:=\varphi(\mathbf{p})$. The vector field is hence defined as a desired velocity vector
\begin{equation}\label{eqq:3}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., it's vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}^{\geq 0}$ the gain which adjusts the speed of convergence of the vector field to the desired trajectory.

@@

Under the assumption of the airspeed $s=\|w(t)\|$ for $t>0$ (i.e., the constant airspeed being greater than the norm of the wind), the guidance action defines how fast the drone converges to the desired trajectory and can be expressed
\begin{equation}
  u(\dot{\mathbf{p}},\mathbf{p},\psi)=\frac{\|\mathbf{p}\|}{s\cos{\beta}}\left( \dot{\mathbf{\chi}}_d(\dot{\mathbf{p}},\mathbf{p})+k_d\hat{\dot{\mathbf{p}}}^TE\hat{\dot{\mathbf{p}}}_d \right),
\end{equation} 
where $k_d\in\mathbb{R}^{\geq 0}$ is the gain which adjusts the speed of convergence of the drone to the vector field $\dot{\mathbf{p}}_d$ to the desired trajectory, $\beta=\cos^{-1}{\left( \hat{\dot{\mathbf{p}}}^T\Psi(\psi) \right)}$ is the slideslip angle, and $\dot{\chi}_d$ is given in Equation~(\ref{eqq:4}).

@@

We consider a Fourier series of an arbitrary order $r$ to evaluate the energy spent performing a given motion and software configuration
\begin{equation}\label{eqq:6}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients. The Fourier analysis allows to account for the periodicity of the mission, with a trajectory $\mathcal{P}$ being reitared over time.

The non-linear model in Equation~(\ref{eqq:6}) can be expressed using an equivalent linear time-varying state-space model, expressed in the following form:
\begin{equation}\label{eqq:7}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)+w(t),\\
  y(t)&=C\mathbf{q}(t)+v(t),
\end{cases}\end{equation}
where $y(t)\in\mathbb{R}^{\geq 0}$ is the energy evolution of the controlled system, $w(t)$ the same process noise from the system in Equation~(\ref{eqq:1}), $v(t)\in\mathbb{R}^{\geq 0}$ the measurement noise, and the control $\mathbf{u}$ along with the input matrix $B$ are defined subsequently. The state $\mathbf{q}$ respresents the evolution of the Fourier series coefficients in time, which can be expressed along the state transition matrix, and the output matrix as follows
\begin{equation}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{ccccccc}
    a_0 & \vline & a_1 & b_1 & \cdots & a_r & b_r
  \end{array}\right],\\
  A&=\left[\begin{array}{ccccc}
    1          & \vline & \mathbf{0} & \cdots & \mathbf{0} \\\hline
    \mathbf{0} & \vline & A_1        & \cdots & \mathbf{0} \\
    \vdots     & \vline & \vdots     & \vdots & \vdots     \\
    \mathbf{0} & \vline & \mathbf{0} & \cdots & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & n \\ -n^2 & 0\end{bmatrix},\\
  C&=\left[\begin{array}{ccccccc}
    1 & \vline & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}\in\mathbb{R}^l$ given $l:=2r+1$, $A\in\mathbb{R}^{l\times l}$ is the state transmission matrix, $C\in\mathbb{R}^l$ is the output matrix respectively. Furthermore, the first row and column of the matrix $A$ contain a zeros row vectors and columns vectors respectively.

Motivated by the fact that the system of interest is sampled at discrete time, and for the sake of simplicity, we consider the discretized version of Equations~(\ref{eqq:1}~and~\ref{eqq:7}) by employing the following expression:
\begin{subequations}\begin{align}
  &\begin{cases}
  \mathbf{p}_{k+1}&=s\Psi(\psi_k)+w_k,\\
  \psi_{k+1}&=u_k,
  \end{cases}\tag{9b}\label{eqq:9b}\\
  &\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+w_k,\\
  y_k&=C\mathbf{q}_k+v_k.
  \end{cases}
\end{align}\end{subequations}

@@

The prediction is done using the following expression:
\begin{subequations}\label{eqq:10}\begin{align}
  \hat{\mathbf{q}}_{k+1}^-&=A\hat{\mathbf{q}}_{k}+B\mathbf{u}_k,\label{eqq:10a}\\
  P_{k+1}^-&=AP_kA^T+Q,\label{eqq:10b}
\end{align}\end{subequations}
where $\hat{\mathbf{q}}_k^-,\hat{\mathbf{q}}_k\in\mathbb{R}^l$ depicts the estimate of the state before and after measurement (or simply estimate), and $P_k,P_k^-\in\mathbb{R}^{l\times l}$ the error covariance matrix (i.e., the variance of the estimate before measurement).

The estimation of the state and the update of the predicted output is done using the following expression:
\begin{subequations}\label{eqq:11}\begin{align}
  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k-C\hat{\mathbf{q}}_{k+1}^-),\label{eqq:11b}\\
  P_{k+1}&=(I-G_{k+1}C)P_{k+1}^-,\\
  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eqq:11d}
\end{align}
\end{subequations}
where $K_k\in\mathbb{R}^l$ is the gain of the Kalman filter, and $I$ the identity matrix. The noise covariance matrixes $Q\in\mathbb{R}^{l\times l},R\in\mathbb{R}$ indicates the process noise and sensor noise covariance respectively, and $\hat{y}_k\in\mathbb{R}$ is the estimated energy.

@@

The energy due to the computational units of the system is obtained with \powprof{}, an open-source modeling tool that measures empirically a number of software configuration and build an energy model accordingly~\cite{seewald2019coarse}. A multivariate linear interpolation is derived automatically by \powprof{}, while being accessed online at the hand of a lookup table in the optimal control algorithm presented later. The robotics system in such an analysis composes a number of computationally expensive ROS nodes, allowing to vary the amount of computations changing node-specific quality of service (QoS) values. For the sake of simplicity and to allow an easy integration in an existing ROS system, the QoS values are utilizing ROS middleware, meaning that defined ROS parameters are intended specifically as QoS ranges.

Let us define $\mathrm{QoS}_i\in\mathbb{Z}^{\geq 0},\,\forall i\in\{0,\dots,r\}$ the $i$-th QoS range, $g_c(\mathrm{QoS}_i)\in\mathbb{R}^{\geq 0}$ the instantaneous energy value obtained interogating \powprof{} online, and $\mathbf{u}_c$ the set of $r$ QoS values the system is composed of. The computational energy component can be hence described using the following expression:
\begin{equation}
  E_c(\mathbf{u}_c)=\sum_{i=0}^{r}{g_c(\mathrm{QoS}_i)},
\end{equation}
TODO

>>>>

%% math package
\usepackage[cmex10]{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{arydshln}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

@@

Let us define a generic continuously differentiable function $\varphi:\mathbb{R}^2\rightarrow\mathbb{R}$. The desired trajectory $\mathcal{P}$ can be hence defined:
\begin{equation}\label{eqq:2}
  \mathcal{P}:=\{\mathbf{p}:\|\varphi(\mathbf{p})\|<c+\varepsilon\},
\end{equation}
as the set of points that follows the trajectory $\varphi(\mathbf{p})=c$ within a given $\varepsilon\in\mathbb{R}_{\geq 0}$.

Given a set of discrete values $c$, the space relative to $\mathcal{O}_W$ can be covered by all the trajectories $\mathcal{P}$ within specific boundaries $\underline{c}\leq c\leq \overline{c}$, where $\underline{a},\overline{a}$ returns the upper bound and lower bound limit of $a$ from the mission specification, which is a lookup table.

The concept of a trajectory is used later to design a controller that selects $c$ with the highest energy value under the energy budget constraints defined in the next subsection. Here we design a guidance action $u$ that allows following such trajectory. The guidance is derived using the vector field approach presented in~\cite{de2017guidance}. The algorithm heads to $\mathcal{P}$ using a vector field based on the minimization of the norm in Equation~(\ref{eqq:2}).

Let us define $\varPhi:=\varphi(\mathbf{p})$. The vector field is hence defined as a desired velocity vector
\begin{equation}\label{eqq:3}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., it's vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain which adjusts the speed of convergence of the vector field to the desired trajectory.

@@

Under the assumption of the airspeed $s=\|w(t)\|$ for $t>0$ (i.e., the constant airspeed being greater than the norm of the wind), the guidance action defines how fast the drone converges to the desired trajectory and can be expressed
\begin{equation}\label{eqq:5}
  u(\dot{\mathbf{p}},\mathbf{p},\psi)=\frac{\|\mathbf{p}\|}{s\cos{\beta}}\left( \dot{\mathbf{\chi}}_d(\dot{\mathbf{p}},\mathbf{p})+k_d\hat{\dot{\mathbf{p}}}^TE\hat{\dot{\mathbf{p}}}_d \right),
\end{equation} 
where $k_d\in\mathbb{R}_{\geq 0}$ is the gain which adjusts the speed of convergence of the drone to the vector field $\dot{\mathbf{p}}_d$ to the desired trajectory, $\beta=\cos^{-1}{\left( \hat{\dot{\mathbf{p}}}^T\Psi(\psi) \right)}$ is the slideslip angle, and $\dot{\chi}_d$ is given in Equation~(\ref{eqq:4}).

@@

We consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ of an arbitrary order $r$ to evaluate the energy spent performing a given motion and software configuration
\begin{equation}\label{eqq:6}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients. The Fourier analysis allows to account for the periodicity of the mission, with a trajectory $\mathcal{P}$ being reitared over time.

The non-linear model in Equation~(\ref{eqq:6}) can be expressed using an equivalent linear time-varying state-space model, expressed in the following form:
\begin{equation}\label{eqq:7}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)+w(t),\\
  y(t)&=C\mathbf{q}(t)+v(t),
\end{cases}\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the energy evolution of the controlled system, $w(t)$ the same process noise from the system in Equation~(\ref{eqq:1}), $v(t)\in\mathbb{R}_{\geq 0}$ the measurement noise, and the control $\mathbf{u}$ along with the input matrix $B$ are defined subsequently. The state $\mathbf{q}$ respresents the evolution of the Fourier series coefficients in time, which can be expressed along the state transition matrix, and the output matrix as follows
\begin{equation}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{ccccccc}
    a_0 & \vline & a_1 & b_1 & \cdots & a_r & b_r
  \end{array}\right],\\
  A&=\left[\begin{array}{ccccc}
    1          & \vline & \mathbf{0} & \cdots & \mathbf{0} \\\hline
    \mathbf{0} & \vline & A_1        & \cdots & \mathbf{0} \\
    \vdots     & \vline & \vdots     & \ddots & \vdots     \\
    \mathbf{0} & \vline & \mathbf{0} & \cdots & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & n \\ -n^2 & 0\end{bmatrix},\\
  C&=\left[\begin{array}{ccccccc}
    1 & \vline & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, $C\in\mathbb{R}^j$ is the output matrix respectively. Furthermore, the first row and column of the matrix $A$ contain a zeros row vectors and columns vectors respectively.

Motivated by the fact that the system of interest is sampled at discrete time, and for the sake of simplicity, we consider the discretized version of Equations~(\ref{eqq:1}~and~\ref{eqq:7}) by employing the following expression
\begin{subequations}\begin{align}
  &\begin{cases}
  \mathbf{p}_{k+1}&=s\Psi(\psi_k)+w_k,\\
  \psi_{k+1}&=u_k,
  \end{cases}\tag{9b}\label{eqq:9b}\\
  &\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+w_k,\\
  y_k&=C\mathbf{q}_k+v_k.
  \end{cases}
\end{align}\end{subequations}

@@

The prediction is done using the following expression:
\begin{subequations}\label{eqq:10}\begin{align}
  \hat{\mathbf{q}}_{k+1}^-&=A\hat{\mathbf{q}}_{k}+B\mathbf{u}_k,\label{eqq:10a}\\
  P_{k+1}^-&=AP_kA^T+Q,\label{eqq:10b}
\end{align}\end{subequations}
where $\hat{\mathbf{q}}_k^-,\hat{\mathbf{q}}_k\in\mathbb{R}^j$ depicts the estimate of the state before and after measurement (or simply estimate), and $P_k,P_k^-\in\mathbb{R}^{j\times j}$ the error covariance matrix (i.e., the variance of the estimate before measurement).

The estimation of the state and the update of the predicted output is done using the following expression:
\begin{subequations}\label{eqq:11}\begin{align}
  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k-C\hat{\mathbf{q}}_{k+1}^-),\label{eqq:11b}\\
  P_{k+1}&=(I-G_{k+1}C)P_{k+1}^-,\\
  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eqq:11d}
\end{align}
\end{subequations}
where $K_k\in\mathbb{R}^j$ is the gain of the Kalman filter, and $I$ the identity matrix. The noise covariance matrixes $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}$ indicates the process noise and sensor noise covariance respectively, and $\hat{y}_k\in\mathbb{R}_{\geq 0}$ is the estimated energy.

@@

A computational energy model is built using \powprof{}\footnote{\label{cross-ref:powprof-link}\url{https://bitbucket.org/adamseew/powprofiler}}, an open-source modeling tool that measures empirically a number of software configuration and build an energy model accordingly~\cite{seewald2019coarse}. The tool builds a multivariate linear interpolation which is accessed online at the hand of a lookup table in the optimal control algorithm. An existing ROS system composes a number of computationally expensive ROS nodes, allowing to vary the amount of computations changing node-specific quality of service (QoS) values. The tool builds the energy model using a file specification which specifies per each ROS node a QoS range.

Suppose the system is composed of $\sigma$ computationally expensive ROS nodes. Let us define the computational control action as
\begin{equation}\label{eqq:12}
  \mathcal{C}_k:=\left\{u:u\in\text{QoS}_n(k)\,\forall\,n\in\{0,\dots,\sigma\}\right\},
\end{equation}
where $\text{QoS}_n(t):\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $n$-th QoS value at time $t$, and $\mathcal{C}_k\in\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ QoS values the system is composed of at time $k$. Let us further define $g:\mathbb{Z}_{\geq 0}\times\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous energy value obtained interogating \powprof{}. The computational energy component can be hence described
\begin{equation}\begin{split}
  g(\mathcal{C}_{k})&:=\{g\left(0,\text{QoS}_0(k)\right),\dots,g\left(\sigma,\text{QoS}_\sigma(k)\right)\},\\
  y_c(t)&=\sum_{x\in g\left(\mathcal{C}_{\ceil{t}}\right)}{x},
\end{split}\end{equation}
where $\ceil{\cdot}$ is the ceiling function that returns the least integer greater than or equal to the argument.

The QoS parameters $\mathcal{C}_k$ can be subject to different constraints at different states. Physically, this means that the drone can perform the ROS nodes within different QoS ranges while flying different phases of a mission
\begin{equation}\label{eqq:14}
  \underline{\text{QoS}}_n(k)\leq \text{QoS}_n(k)\leq \overline{\text{QoS}}_n(k),\,\forall n\in\{0,\cdots,\sigma\},
\end{equation}
where the values $\underline{\text{QoS}}_n(k),\overline{\text{QoS}}_n(k)$ are retrived from the mission specification.

The control action is constructed in two steps. Equation~(\ref{eqq:12}) defines the control due to the computational model. The control due to the energy model, which describes the energy of the mechanical elements of the drone $\mathcal{M}$, is derived in the following subsection.

\subsection{Control Action}

Given a generic trajectory equation $\varphi$, the trajectory can be modeled by $\rho$ parameters $\mathbf{u}\in\mathbb{R}^\rho$, e.g., the constants of a linear function, the radius of a circonference, and semi-major and minor axis of an ellipse. The set of these parameters can be expressed
\begin{equation}\label{eqq:14}
  \mathcal{M}_k:=\{\mathbf{u}_k : \varphi_k(\mathbf{p}_k^0,\mathbf{u}_k)=c_k\},
\end{equation}
where $c$ is being defined in Equation~(\ref{eqq:2}), and $\mathbf{p}^0$ is the optimal point which lets the trajectory explicit function $\varphi_k$ converge to the value $c$.

The explicit trajectory equation $\varphi_k$ can be different for different states $k$, meaning the vector field and guidance action, from Equations~(\ref{eqq:3})~and~(\ref{eqq:5}) respectively, will account for the sudden change of trajectory during the mission. Alike Equation~(\ref{eqq:14}), the parameters $\mathbf{u}_k=\{u_{k,1},\dots,u_{k,\rho_k}\}$ at time $k$ are constrained
\begin{equation}
  \underline{u}_{k,n}\leq u_{k,n}\leq \overline{u}_{k,n}\,\forall n\in\{0,\cdots,\rho_k\},
\end{equation}
where the values $\underline{u}_{k,n},\overline{u}_{k,n}$ are also retrived from the mission specifcation.

It is worth considering that the number of parameters at state $k$ is also a parameter of the state. This is for the sake of generality, as the mission specification might contain different explicit equations for different states. Meaning that the drone might follow in an ellipse function thorought the mission and heading a linear function while landing.

The mechanical and computational control actions, $\mathcal{C}$ and $\mathcal{M}$ defined in Equation~(\ref{eqq:12})~and~(\ref{eqq:14}) respectively, are incorporated in the system in Equation~(\ref{eqq:9b}) using the input matrix
\begin{equation}\begin{split}
  \mathbf{u}_k&=\left[\begin{array}{cc}
    g\left(\mathcal{C}_k\right) & \mathcal{M}_k
  \end{array}\right]^T,\\
  B&=\left[\begin{array}{ccccccc}
    1 & \cdots & 1 & \vline & 0 & \cdots & 0\\\hline
    0 & \cdots & 0 & \vline & 0 & \cdots & 0\\
    \vdots & \ddots & \vdots & \vline & \vdots & \ddots & \vdots\\
    0 & \cdots & 0 & \vline & 0 & \cdots & 0
  \end{array}\right],
\end{split}
\end{equation}
where $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=\sigma+\rho$, and $B\in\mathbb{R}^{j\times l}$ is the input matrix from Equation~(\ref{eqq:9b}). The first $\sigma$ columns in the first row of the input matrix are $1$, while all the other items are $0$. This adds the computational model component to the energy evolution in the system in Equation~(\ref{eqq:9b}). The energy due to the change of explicit trajectory equation parameters is not directly added to the system, which hovewever will update the reading from the sensors in Equation~(\ref{eqq:11b}) and thus adjust the energy evolution accordingly.
